---
title: "Untitled"
author: "Shankar K. Shakya"
date: "11/4/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Reading data file first

```{r}

mydata <- read.csv("fungicide_dat.csv")
head(mydata)

```

## How many rows and columns are in this dataset?


## Convert bu/acre to kg/ha and add this as new column
This can be done by multiplying bu/acre with 62.77. 

```{r}

mydata$kg_ha <- mydata$Yield_bu_per_acre * 62.77
head(mydata)

```

## subset your data only to select your treament and kg_ha column

```{r}

yield_kg_ha <- mydata[ , c(1,4)]
head(yield_kg_ha)

```

## Some statistical analysis

```{r}

?aov

myanova <- aov(formula = Yield_bu_per_acre ~ Treatment, data= mydata) 

myanova


summary(myanova)

TukeyHSD(myanova) 

```


## Please do similar analysis for severity to test for statistical significance


## Now lets visualize our data

```{r}
library(ggplot2)

yield.plot <- ggplot(data = mydata, aes(x = Treatment, 
                                    y = Yield_bu_per_acre))
yield.plot


```


```{r}
yield.plot <- yield.plot + geom_boxplot()

yield.plot
```


```{r}

yield.plot <- yield.plot + geom_boxplot(fill = "orange")
yield.plot

```

```{r}

yield.plot <- yield.plot + geom_boxplot(aes(fill = Treatment))
yield.plot

```


```{r}
yield.plot <- yield.plot + ggtitle("Effect of Fungicides on Yield")
yield.plot

```

```{r}
yield.plot <- yield.plot + xlab("Treatment Applied") + ylab("Yield")
yield.plot
```


## Now try plotting disease severity data 


```{r}
ggplot(mydata, aes(x = Treatment, y = Severity)) + geom_boxplot(aes(fill = Treatment))

severity.plot <- ggplot(mydata, aes(x = Treatment, y = Severity)) + geom_boxplot(aes(fill = Treatment)) + theme_bw(base_size = 14) + theme(panel.grid = element_blank())

severity.plot

```


```{r}
library(ggpubr)
severity.plot <- severity.plot + stat_compare_means(method = "anova", label.y = 6.5)

severity.plot
```



```{r}
my_comparisons <- list(c("Control", "Fungicide_A"),
                       c("Control", "Fungicide_B"), 
                       c("Fungicide_A", "Fungicide_B"))

severity.plot <- severity.plot +
    stat_compare_means(comparisons = my_comparisons, 
                        label = "p.signif", 
                        method = "t.test")

severity.plot

```


## Now we are ready to save our plot

```{r}

#tiff("myfungicidedata.tiff", width = 7, height = 7, res = 300, units = "in", compression = "zip")
severity.plot
#dev.off()


```


```{r}
#tiff("myfungicidedata.tiff", width = 9, height = 7, res = 300, units = "in", compression = "zip")
severity.plot
#dev.off()
```


