---

---

## What is R and why R?

R is a free statistical programming language. It has been widely used for statistics, data analysis and data visualization. One of the fields where R is widely used is genomics. Many R packages have been developed for population genetics data analyses and visualization.

### Installing R and Rstudio

Download and install [R](http://cran.r-project.org/) and [RStudio](http://www.rstudio.com/). 
if you do not have them already. Rstudio is optional but it is a great tool if you are just starting to learn R.
Once you have download and installed R and Rstudio, we ll start working by creating project.


### R as calculator

```{r}
# Add

5+10

# Substract

10-5 

# Multiply
5 * 10


# Divide
100 / 5


```

### R data structures

```{r}

# vectors
x <- c(1,3,2,10,5)
x

# Matrices
x <- c(1,2,3,4)
y <- c(4,5,6,7)
matrix <- cbind(x,y)
mymat <- matrix(NA, nrow = 2, ncol = 2)

# Data frame
chr <- c("chr1", "chr1", "chr2", "chr2")
strand <- c("-","-","+","+")
start<- c(200,4000,100,400)
end<-c(250,410,200,450)
mydata <- data.frame(chr,start,end,strand)
#change column names
names(mydata) <- c("chr","start","end","strand")
mydata 

# List

w <- list(name="Fred",
       mynumbers=c(1,2,3),
       mymatrix=matrix(1:4,ncol=2),
       age=5.3)
w

```



