---

---

```{r setu, include=FALSE}
knitr::opts_chunk$set(fig.width = 9, fig.height = 7, echo = T, warning = F, message = F, verbose = F)
```

# Analyses of microsatellite data in R

Hopefully I have covered what microsatellites markers are by this time. Many R packages are available to analyse microsatellite data. We ll be using R package Poppr to analyze the data from [Shakya et al. 2018](https://apsjournals.apsnet.org/doi/10.1094/PDIS-11-17-1801-RE). 

We ll first read the paper and understand what were the objectives of the paper and try to reproduce the work. For this part, We ll work in a group of 3-4 students.

Lets remind ourselves what these terms mean: Diploid individuals, multilocus genotypes, locus/loci, diversity index, index of association, rbarD, Hardy Weinberg Equilibrium, Fst, population structure, DAPC, neighbor joining tree and Mantel's test.


# Download the necessary data

First, we need the microsatellite genotype data which can be downloaded from [here](https://osf.io/hbxvj/files/). Download MX_pop.csv file to your current working directory. 


```{r}

library(poppr)
P.inf <- read.genalex("MX_pop.csv", ploidy = 4)
P.inf

```

## Lets look at the missing data

```{r, fig.width = 12, fig.height = 6, fig.align="center"}
info_table(P.inf, type = "missing", plot = T)
```


Now instead of working with tetraploids lets work with diploid data.

```{r, fig.width = 12, fig.height = 6, fig.align="center"}

multiploid2diploid <- function(x, to = 2){
  
  ploidy_tab <- info_table(x, type = "ploidy")
  mat_sum <- rowSums(ploidy_tab > to, na.rm = TRUE)
  
  diploid_pinf <- x[mat_sum == 0]
  
  diploid_pinf <- recode_polyploids(diploid_pinf, newploidy = to)
  
  return(diploid_pinf)
  
}


diploid_pinf <- multiploid2diploid(P.inf, to = 2)
diploid_pinf

info_table(diploid_pinf, type = "missing", plot = T)

```


## Remove locus with high percentage of missing data

```{r}

diploid_pinf <- missingno(diploid_pinf, type = "loci", cutoff = 0.07)
splitStrata(diploid_pinf) <- ~Region/Year
setPop(diploid_pinf) <- ~Region

diploid_pinf

```

## Table 2. Diversity summary by region 

```{r}
# Corrected simpson diversity index

uSimp <- function(x){
  lambda <- vegan::diversity(x, "simpson")
  x <- drop(as.matrix(x))
  if (length(dim(x)) > 1){
    N <- rowSums(x)
  } else {
    N <- sum(x)
  }
  return((N/(N-1))*lambda)
}


unbiased <- poppr(diploid_pinf, uSimp = uSimp)
unbiased <- unbiased[, c(1,2,3,4,9,10,11)]
colnames(unbiased) <- c("Region", "N", "MLG", "eMLG", "E5", "Corrected lambda", "Hexp")
unbiased <- unbiased[c(6,1,4,3,2,5), ]
rownames(unbiased) <- NULL

unbiased

```

## Table 3. rbarD and Hardy-Weinberg equilibrium

```{r, fig.width = 9, fig.height = 7, fig.align="center"}

ia_Pinf <- poppr(diploid_pinf, clonecorrect = TRUE, strata = ~Region, sample = 999, quiet = TRUE)

ia_Pinf <- ia_Pinf[,c(1:3,13:14)]  
colnames(ia_Pinf) <- c("Region", "N", "MLG", "rbarD", "P-value")
ia_Pinf

```

## HWE

```{r, fig.width = 9, fig.height = 7, fig.align="center"}

library(pegas)
diploid_pinf_cc <- clonecorrect(diploid_pinf)
hwe_per_pop <- seppop(diploid_pinf_cc) %>% lapply(hw.test, B = 1000)
per_pop_mat <- sapply(hwe_per_pop, "[", i = TRUE, j = 3)
alpha <- 0.05
per_pop_mat[per_pop_mat > alpha] <- 1
library(lattice)
hwe_plot <- levelplot(per_pop_mat, xlab = "Locus", ylab = "Population")
num_loci_hwe <- per_pop_mat
num_loci_hwe[num_loci_hwe != 1] <- 0
num_loci_hwe <- colSums(num_loci_hwe)
ia_Pinf <- ia_Pinf[-7, ]
ia_Pinf <- cbind(ia_Pinf, num_loci_hwe) 
ia_Pinf <- ia_Pinf[c(6,1,4,3,2,5), ]
rownames(ia_Pinf) <- NULL
colnames(ia_Pinf)[6] <- "Loci under HWE"

ia_Pinf

```

## Table 4. Pairwise FST 

```{r, message=FALSE, warning=FALSE, fig.width = 9, fig.height = 7, fig.align="center"}

diploid_pinf_cc <- clonecorrect(diploid_pinf)

library(strataG)
gen2gtype <- function (x) {
  
  gen.mat <- genind2df(x, usepop = TRUE, oneColPerAll = TRUE)
  strata <- x@strata
  rownames(strata) <- rownames(gen.mat)
  x@strata <- strata
  gen.mat[gen.mat == "NA"] <- NA
  has.pop <- !is.null(x@pop)
  df2gtypes(x = gen.mat, ploidy = x@ploidy[1], id.col = NULL, 
            strata.col = if (has.pop) 
              1
            else NULL, loc.col = if (has.pop) 
              2
            else 1, 
            schemes = x@strata,  other = other(x))
}

diploid_pinf_cc.g <- gen2gtype(diploid_pinf_cc)

pairwise_fst <- pairwiseTest(diploid_pinf_cc.g, stats = "fst", nrep = 100)

fst_mat <- pairwise_fst$pair.mat$Fst
fst_mat[upper.tri(fst_mat)] <- t(fst_mat)[upper.tri(fst_mat)]
fst_mat[is.na(fst_mat)] <- 0


library(reshape2)
fst_mat[upper.tri(fst_mat)] <- NA
lower_tri <- melt(fst_mat, na.rm = TRUE)


library(ggplot2)
ggheatmap <- ggplot(lower_tri, aes(x = Var1, y = Var2, fill = value)) + geom_tile(color = "white") +
  scale_fill_gradient(low = "green", high = "red" , space = "Lab", name="Pairwise FST") + theme_minimal() +
   theme(axis.text.x = element_text(angle = 90, vjust = 1, size = 12, hjust = 1)) + coord_fixed() +
  labs(x = "Population", y = "Population") +
  theme(axis.text.y = element_text(size = 12)) +
  theme(axis.title = element_text(size = 12)) +
  geom_text(aes(label = round(value, 2)))

ggheatmap

```


## Table 5. Analysis of molecular variance

```{r}

repeat_length <- c(2, 2, 2, 3, 3, 2, 2, 2, 2, 2, 2)
bruvodist <- bruvo.dist(clonecorrect(diploid_pinf), replen = repeat_length)

amv <- poppr.amova(diploid_pinf, 
                   hier = ~Region, 
                   clonecorrect = TRUE, 
                   dist = bruvodist,
                   within = F)


myamv <- amv$results
rownames(myamv) <- c("Between regions", "Within regions", "Total")
myamv <- cbind(myamv, amv$componentsofcovariance$`%`)

colnames(myamv) <- c("df", "SS", "MSS", "% variance")

myamv

```


## Fig 2. Neighbor joining tree

```{r, fig.width = 12, fig.height = 6, fig.align="center"}

repeat_length <- c(2, 2, 2, 3, 3, 2, 2, 2, 2, 2, 2)

diploid_pinf <- clonecorrect(diploid_pinf)
tree_nj <- bruvo.boot(diploid_pinf, replen = repeat_length, 
                   sample = 99, cutoff = 70, tree = "nj")

plot(tree_nj)




```

## STRUCTURE analysis

```{r, fig.width = 12, fig.height = 6, fig.align="center"}

edited_qmat_list <- readRDS("qmat_list.RData")

for (i in 1:length(edited_qmat_list)){
  temp_mat <- edited_qmat_list[[i]]
  key <- c("Michoacan" = "MICHO", "Chapingo" = "CHA", "Toluca" = "TOLU", "San Gerónimo" = "SG", "Juchitepec" = "JFH", "Tlaxcala" = "TLAX")
  temp_mat$orig.pop <- factor(key[temp_mat$orig.pop], ordered = TRUE, levels = key)
  edited_qmat_list[[i]] <- temp_mat
} 

library(strataG)
structurePlot(edited_qmat_list[[3]], horiz = F)

```


## Fig 5. Discriminant Analysis of Principal components

```{r, eval = F, fig.width = 12, fig.height = 6, fig.align="center"}

set.seed(999)
P.infx <- xvalDapc(tab(diploid_pinf, NA.method = "mean"), pop(diploid_pinf),
                   n.pca = 60:70,  n.rep = 100, 
                   parallel = "multicore", ncpus = 4L)

library(RColorBrewer)
myCol <- brewer.pal(nPop(diploid_pinf), "Dark2") %>% setNames(popNames(diploid_pinf))

diploid_pinf_no_MCH <- popsub(diploid_pinf, blacklist = "Michoacan")


set.seed(999)
P.infx1 <- xvalDapc(tab(diploid_pinf_no_MCH, NA.method = "mean"), pop(diploid_pinf_no_MCH),
                   n.pca = 60:70,  n.rep = 100,
                   parallel = "multicore", ncpus = 4L)

source("scatterdapc.R")

par(mfrow = c(1,2))
par(mar = c(1,1,1,3))

scatterdapc(P.infx$DAPC, col=myCol, clabel = 0.75, pch=15:19, scree.pca = TRUE, scree.da = FALSE, 
        posi.pca = "topright", posi.leg = "topright", legend = F, 
        cleg = 0.9, inset.solid = 1, xax = 1, yax = 2, cex.lab = 1, cex = 1, solid = 1, cstar = 0)
title(main = "A", col = "black", font.main = 2, adj = 0, line = -1)

scatterdapc(P.infx1$DAPC, col=myCol, clabel = 0.75, pch=15:19, scree.pca = TRUE, scree.da = FALSE, 
        posi.pca = "topright", posi.leg = "topright", legend = F, 
        cleg = 0.9, inset.solid = 1, xax = 1, yax = 2, cex.lab = 1, cex = 1, solid = 1, cstar = 0)
title(main = "B", col = "black", font.main =2, adj = 0, line = -1)



```


